---
title: "Project01_IB"
author: "Ian Beren"
date: '2022-10-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

dat <- read.csv('lead-iq-01.csv')
library(ggplot2)
library(sqldf)
library(tidyverse)
```




```{r}
ggplot(data=dat, aes(Smelter, IQ)) + stat_summary(geom='bar', fun='mean')

meanvars = aggregate(dat$IQ, list(dat$Smelter), FUN=mean)
```
Based on the graph, we can see the mean for those who live further from the smelter is higher (Mean = `r round(meanvars[1,2],2)`) then for those who live nearer to the smelter (Mean = `r round(meanvars[2,2],2)`). 

```{r}
#is.data.frame(dat)


knitr::kable (dat %>%                               
  group_by(Smelter) %>% 
  summarize(Mean = mean(IQ),
            StdDev = sd(IQ),
            Minimum = min(IQ),
            Q1 = quantile(IQ, 0.25),
            Median = median(IQ),
            Q3 = quantile(IQ, 0.75),
            Maximum = max(IQ)), digits=2)
```

Based on the summary statistics provided above, it is noted there there is a potential data issue with the maximum value of IQ for those who live further from the smelter, having a value of 999. This will likely need to be addresed before further analysis is conducted.
